---
title: "Supply Chain Risk and Profitability Analysis"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

# Introduction

## Objective
In modern supply chain management, late deliveries and unprofitable orders represent major operational losses. In technology companies an executive report stipulates that technology companies lose on average $1 million per logistics disruption/incident. This then leads to customer complaints, lost business and contracts. When supply chain disruptions occur on a large scale, restocking delays and high storage costs can lead to stockouts especially for goods intensive for delayed inputs. This project aims to analyze the extent at which both issues arise by building predictive models on historical logistics data. We will pursue and research these two primary learning tasks:

1. Classification (Supervised): Determine what predictors influence whether a shipment is at risk of delayed delivery (the binary target variable "Late_delivery_risk").
2. Regression (Supervised): Forecast the exact profit margin of an order ("Benefit per order") based on checkout-time transaction details.

```{r setupkaggle, include=FALSE, echo=FALSE}
library(RKaggle)
library(rstatix)
library(dplyr)
library(zeallot)
library(purrr)
library(lubridate)
library(ggplot2)
library(tidyr)
library(forcats)
```

To get RKaggle to pull dataset automatically:
1. Go to "Your API Tokens" on Kaggle.com and hit generate
2. Copy + put api-key generated on website into `~/.kaggle/kaggle.json`
3. Make sure RSession is restarted
4. Then run cell


```{r dataset}
# Get dataset and download once
if (exists("dataset")==0) {
  c(dataset, descriptions, data_staleness_log) %<-% get_dataset("shashwatwork/dataco-smart-supply-chain-for-big-data-analysis")
}
dataset <- dataset %>%
  mutate(across(where(is.character), ~ iconv(.x, from = "ISO-8859-1", to = "UTF-8")))

# dataset <- read.csv("DataCo_cleaned.csv")
```


```{r summary, echo=FALSE} 
summary(dataset)
```
Dataset has a size of 180k rows and 53 features but only about 45 are useful (customer email, fname, lname and street are not super useful) and 2 are predictor variables

```{r size}

targets = c("Late_delivery_risk", "Benefit per order")
dataset %>% dim()
dataset %>% names()
dataset %>% str()
```
Late_delivery_risk: "Categorical variable that indicates if sending is late (1), it is not late (0)."

Benefit per order: "Earnings per order placed"

**Candidate Predictors** would be Sales, Order Region (or some location for order), Product Category ID, Product Price, Customer Region (or some location for customer), Days for shipping (real), Shipping mode, Delivery Status, to name a few


Data Cleaning

Step 1: Missing value

```{r}
# Report missing counts and percentages
missing_summary <- data.frame(
  Variable = names(dataset),
  Missing_Count = sapply(dataset, function(x) {
    sum(is.na(x) | trimws(as.character(x)) == "")
  }),
  Missing_Percent = sapply(dataset, function(x) {
    mean(is.na(x) | trimws(as.character(x)) == "") * 100
  })
)

missing_summary %>%
  arrange(desc(Missing_Percent))
```
```{r}
df <- dataset %>%
  select(-any_of(c(
    "Product Description",
    "Order Zipcode"
  )))
```
### Step 2: Duplicate:
```{r}
# Count exact duplicate rows
sum(duplicated(df))

# Inspect duplicate rows
df[duplicated(df), ]

# Remove exact duplicates
df <- df %>%
  distinct()
```
#### Step 3: Data Types
```{r}
df <- df %>%
  mutate(
    `Late_delivery_risk` = as.factor(Late_delivery_risk),
    `Shipping Mode` = as.factor(`Shipping Mode`),
    `Customer Segment` = as.factor(`Customer Segment`),
    `Market` = as.factor(Market),
    `Order Region` = as.factor(`Order Region`),
    `Days for shipment (scheduled)` = as.numeric(`Days for shipment (scheduled)`),
    `Benefit per order` = as.numeric(`Benefit per order`),
    order_date = mdy_hm(`order date (DateOrders)`
    )
  )

str(df)
```
### Step 4. Unnecessary Variables
```{r}
unnecessary_vars <- c(  "Customer Email", "Customer Password", "Customer Fname", "Customer Lname", "Customer Street", "Product Image", "Product Description", "Product Status", "Delivery Status", "Days for shipping (real)")

df_clean <- df %>%
  select(-any_of(unnecessary_vars))
head(df_clean)
```

## Early Observations

* Orders date between [2015,2018]
* ~20k customers purchasing ~65k products from 563 cities from the United States and Puerto Rico

```{r}
df_clean %>% 
  mutate(year = year(order_date)) %>% 
  count(year)

paste("Distinct customers: ", df_clean %>% select(`Customer Id`) %>% distinct %>% count())
paste("Distinct orders: ", df_clean %>% select(`Order Id`) %>% distinct %>% count())
paste("Distinct products: ", df_clean %>% select(`Product Card Id`) %>% distinct %>% count())
paste("Distinct Customer Cities: ", df_clean %>% select(`Customer City`) %>% distinct %>% count())
paste("Distinct Customer Countires: ", df_clean %>% select(`Customer Country`) %>% distinct %>% count())
paste("Distinct Sites: ", df_clean %>% select(Latitude, Longitude) %>% distinct %>% count())
```

## EDA 
### Late_delivery_risk

```{r 3}
top_ten <- c("Days for shipment (scheduled)","Sales per customer", "Latitude", "Longitude", "Order Item Discount", 
  "Order Item Discount Rate", "Order Item Product Price", "Order Item Profit Ratio", "Order Item Quantity", "Order Item Total")

top_ten_cat <- c("Category Name", "Customer City", "Customer Country", "Customer Segment", "Department Name", "Market", "Order City", "Order Country", "Order Region", "Order State", "Product Name", "Shipping Mode")

# Not super informative other than showing how there's a lot of structure in the data
# df_clean %>% 
#   select(
#   `Days for shipment (scheduled)`,`Sales per customer`, Latitude, Longitude, `Order Item Discount`, 
#   `Order Item Discount Rate`, `Order Item Product Price`, `Order Item Profit Ratio`, `Order Item Quantity`, `Order Item Total`) %>% 
#   as.matrix() %>% 
#   pairs(cex=0.1)
# View(df_clean)


df_clean %>% 
  select(top_ten) %>% 
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value") %>%
  ggplot(aes(x = value)) +
  geom_histogram(bins = 20, fill = "royalblue", color = "white") +
  
  # 4. Create separate panels for each variable
  facet_wrap(~ variable, scales = "free") +
  theme_minimal()

df_clean %>% 
  select(`Benefit per order`) %>% 
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value") %>%
  ggplot(aes(x = value)) +
  geom_histogram(bins = 100, fill = "royalblue", color = "white") +
  coord_cartesian(xlim = c(-500, 500)) +
  # 4. Create separate panels for each variable
  facet_wrap(~ variable, scales = "free") +
  theme_minimal()

df_clean %>% 
  ggplot(aes(x = Late_delivery_risk, fill = Late_delivery_risk)) +
  geom_bar( fill = "royalblue", color = "white") +
  labs(x = "Late Delivery Risk", y = "Count") +
  geom_text(stat = "count", 
            aes(group = 1, label = paste0(after_stat(count), " (", round(after_stat(prop), 3)*100, "%)")), vjust = -0.5, size = 4) +

  # coord_cartesian(xlim = c(-500, 500)) +
  theme_minimal()

df_clean %>% 
  select(top_ten_cat, Late_delivery_risk) %>% 
  pivot_longer(cols = top_ten_cat, names_to = "category_name", values_to = "category_value") %>% 
  # Count the pairs first to get the breakdown
  count(category_name, category_value, Late_delivery_risk) %>% 
  # Group by category to find the total volume for slicing
  group_by(category_name, category_value) %>% 
  mutate(total_n = sum(n)) %>% 
  group_by(category_name) %>% 
  # Slice based on the total category size, keeping all risk levels
  slice_max(order_by = total_n, n = 10, with_ties = FALSE) %>% 
  ungroup() %>% 
  mutate(category_value = fct_reorder(category_value, n, .desc = TRUE)) %>% 
  ggplot(aes(x = category_value, y = n, fill = Late_delivery_risk)) + 
  geom_col(color = "white", alpha = 0.8) + 
  facet_wrap(~ category_name, scales = "free", ncol = 2)

```

```{r}
df_clean %>%
  pivot_longer(cols = top_ten, names_to = "predictor", values_to = "x_value") %>%
  ggplot(aes(x = x_value, fill = Late_delivery_risk)) +
  geom_histogram(bins = 20, position = "stack", alpha = 0.7, color = "black") +
  facet_wrap(~ predictor, scales = "free_x", ncol = 4) + # 10 distinct plots
  theme_minimal()

```
### Benefit per Order

```{r 3}

df_clean %>% 
  select(top_ten_cat, `Benefit per order`) %>% 
  pivot_longer(cols = top_ten_cat, names_to = "category_name", values_to = "category_value") %>% 
  # Group by category to find the total volume for slicing
  group_by(category_name, category_value) %>%
  summarise(
    n = n(),
    avg_benefit = mean(`Benefit per order`, na.rm = TRUE),
    total_benefit = sum(`Benefit per order`, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  group_by(category_name) %>%
  slice_max(order_by = n, n = 10, with_ties = FALSE) %>%
  ungroup() %>%
  mutate(category_value = fct_reorder(category_value, avg_benefit, .desc = TRUE)) %>% 
  ggplot(aes(x = category_value, y = n, fill = avg_benefit)) + 
  geom_col(color = "white", alpha = 0.8) + 
  facet_wrap(~ category_name, scales = "free", ncol = 2)


```

```{r}
df_clean %>%
  mutate(
    benefit_per_order_group = cut(
      `Benefit per order`,
      breaks = c(-Inf, -100, 0, 100, Inf),
      labels = c("Very Low", "Low", "High", "Very High")
    )
  ) %>%
  pivot_longer(cols = top_ten, names_to = "predictor", values_to = "x_value") %>%
  ggplot(aes(x = x_value, fill = benefit_per_order_group)) +
  geom_histogram(bins = 20, position = "stack", alpha = 0.7, color = "black") +
  facet_wrap(~ predictor, scales = "free_x", ncol = 4) +
  theme_minimal()

```

### Finding correlations

```{r}
df_clean %>% 
  select(Late_delivery_risk, `Benefit per order`, top_ten) %>% 
  mutate(Late_delivery_risk = as.numeric(Late_delivery_risk)) %>% 
  cor_mat
```

```{r}
df_clean %>% filter(`Product Price` > 1000) %>% 
  select(Late_delivery_risk, `Benefit per order`, top_ten) %>% 
  mutate(Late_delivery_risk = as.numeric(Late_delivery_risk)) %>% 
  cor_mat
```

### Classification Regression

```{r}

df_clean %>% summarise(across(everything(), n_distinct))
usable_cat = c("Category Name", "Customer State", "Department Name", "Market", "Order Region", "Shipping Mode")
predictors = c(top_ten, top_ten_cat)
safe_top_ten <- paste0("`", usable_cat, "`")
lm.fit <- glm(reformulate(safe_top_ten, response="Late_delivery_risk"), data=df_clean, family = binomial(link = "logit"))
summary(lm.fit)
```

