---
title: "Supply Chain Risk and Profitability Analysis"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

# Introduction

## Objective
In modern supply chain management, late deliveries and unprofitable orders represent major operational losses. In technology companies an executive report stipulates that technology companies lose on average $1 million per logistics disruption/incident. This then leads to customer complaints, lost business and contracts. When supply chain disruptions occur on a large scale, restocking delays and high storage costs can lead to stockouts especially for goods intensive for delayed inputs. This project aims to analyze the extent at which both issues arise by building predictive models on historical logistics data. We will pursue and research these two primary learning tasks:

1. Classification (Supervised): Determine what predictors influence whether a shipment is at risk of delayed delivery (the binary target variable "Late_delivery_risk").
2. Regression (Supervised): Forecast the exact profit margin of an order ("Benefit per order") based on checkout-time transaction details.

```{r setupkaggle, include=FALSE, echo=FALSE}
install.packages("RKaggle")
library(RKaggle)
library(dplyr)
library(zeallot)
library(purrr)
```

To get RKaggle to pull dataset automatically:
1. Go to "Your API Tokens" on Kaggle.com and hit generate
2. Copy + put api-key generated on website into `~/.kaggle/kaggle.json`
3. Make sure RSession is restarted
4. Then run cell


```{r dataset}
# Get dataset and download once
if (exists("dataset")==0) {
  c(dataset, descriptions, data_staleness_log) %<-%   get_dataset("shashwatwork/dataco-smart-supply-chain-for-big-data-analysis")
}
```


```{r summary, echo=FALSE} 
summary(dataset)
```
Dataset has a size of 180k rows and 53 features but only about 45 are useful (customer email, fname, lname and street are not super useful) and 2 are predictor variables

```{r size}

targets = c("Late_delivery_risk", "Benefit per order")

dataset %>% dim()
dataset %>% names()
dataset %>% str()
```
Late_delivery_risk: "Categorical variable that indicates if sending is late (1), it is not late (0)."

Benefit per order: "Earnings per order placed"

**Candidate Predictors** would be Sales, Order Region (or some location for order), Product Category ID, Product Price, Customer Region (or some location for customer), Days for shipping (real), Shipping mode, Delivery Status, to name a few


Data Cleaning

Step 1: Missing value

```{r}
# Report missing counts and percentages
missing_summary <- data.frame(
  Variable = names(dataset),
  Missing_Count = sapply(dataset, function(x) {
    sum(is.na(x) | trimws(as.character(x)) == "")
  }),
  Missing_Percent = sapply(dataset, function(x) {
    mean(is.na(x) | trimws(as.character(x)) == "") * 100
  })
)

missing_summary %>%
  arrange(desc(Missing_Percent))
```
```{r}
df <- dataset %>%
  select(-any_of(c(
    "Product Description",
    "Order Zipcode"
  )))
```
### Step 2: Duplicate:
```{r}
# Count exact duplicate rows
sum(duplicated(df))

# Inspect duplicate rows
df[duplicated(df), ]

# Remove exact duplicates
df <- df %>%
  distinct()
```
#### Step 3: Data Types
```{r}
df <- df %>%
  mutate(
    `Late_delivery_risk` = as.factor(Late_delivery_risk),
    `Shipping Mode` = as.factor(`Shipping Mode`),
    `Customer Segment` = as.factor(`Customer Segment`),
    `Market` = as.factor(Market),
    `Order Region` = as.factor(`Order Region`),
    `Days for shipment (scheduled)` = as.numeric(`Days for shipment (scheduled)`),
    `Benefit per order` = as.numeric(`Benefit per order`),
    order_date = mdy_hm(`order date (DateOrders)`
    )
  )

str(df)
```
### Step 4. Unnecessary Variables
```{r}
unnecessary_vars <- c(  "Customer Email", "Customer Password", "Customer Fname", "Customer Lname", "Customer Street", "Product Image", "Product Description", "Product Status", "Delivery Status", "Days for shipping (real)")

df_clean <- df %>%
  select(-any_of(unnecessary_vars))
head(df_clean)
```


## EDA
```{r 3}


```

